### 2016-09-27 Chapters/SolvationAndCorrelation.tex
Luca's comments and corrections are in for the main body of the thesis

@@

It is thus convenient to devise approximations that are able to simplify the
Lagrangian and the governing equations by breaking or weakening the coupling.
While Eq. \eqref{eq:pted-cc-polarization} is \emph{directly} coupled to
the multipliers equation, since the source terms directly depends on
amplitudes and multipliers, it is only \emph{indirectly} coupled to the
amplitudes equation, where only the polarization enters.
Moreover, the leading terms in the correlated source operator
expectation value prominently involve the
singles cluster operator and the singles multiplier state.
Three approximation have currently been proposed: \gls{PTES}, \gls{PTE(S)}
and \gls{PTE}.

@@

To break the coupling, we reformulate the split the system of equations

@@

term. However, this adds another $O(N^5)$ step to the iterative

@@

The \acrshort{PTES} approximation can be brought one step further, by
truncating the \acrshort{CCS} expectation value to its multiplier-independent
term already in the effective Lagrangian.
The coupling between the amplitudes and multipliers is broken, while
still improving upon the reference reaction field with correlated
contributions.\autocite{Caricato2011-tx}
The effective Lagrangian would then read as:
\begin{equation}\label{eq:pte-s-cc}
    {}^{\mathrm{PTE(S)}}\lag(\tamp{}, \tbar{}, \p)_\mathcal{M} =
  {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M}
  + \frac{1}{2}{}^t\p\V\p + {}^t\p\braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
\end{equation}
which uncouples all equations:
\begin{subequations}
  \begin{align}
  {}^{\mathrm{PTE(S)}}\tampEq{\mu_q}(\tamp{}, \tbar{})  &=
   \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
   = 0 \\
   {}^{\mathrm{PTE(S)}}\tbarEq{\mu_q}(\tamp{}, \tbar{}, \p) &=
    \denom{\mu_q}\tbar{\mu_q} +
    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}} +
    \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
    \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}}
    \nonumber \\
    &+
    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
    = 0 \\
    {}^{\mathrm{PTE(S)}}\Omega_\p(\tamp{}, \tbar{}, \p)
    &=
    \V\p + \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}} = 0
    \label{eq:pte-s-polarization}
  \end{align}
\end{subequations}
The amplitude equations are formally equivalent to those derived in the
\acrshort{PTE} scheme, \emph{vide infra}. The equations for multipliers are also formally
equivalent to the \acrshort{PTE} equations, apart from the equation
determining the singles multipliers which features an additional term.
This term is not, however, coupled to the other equations given that
$\p$ is fixed once the singles amplitudes have been determined.

@@

Eventually, the fifth order energy correction is formed including second order
amplitudes, multipliers and polarization:
\begin{equation}\label{eq:pted-mbpt-5th}
  \begin{aligned}
    G^{[5]} &=
    \braket{\mathrm{HF} | \frac{1}{2}
    \commutator{\commutator{\Phi}{T^{[2]}}}{T^{[2]}}
    | \mathrm{HF}}
    +
    \sum_{u=1}^2\braket{\tbar{u}^{[1]}|
    \commutator{\commutator{\Phi}{T^{[2]}}}{T^{[1]}}
    | \mathrm{HF}}
    \\
    &+
    \sum_{u=1}^3\braket{\tbar{u}^{[2]}|
      \commutator{\Phi}{T^{[2]}}
    | \mathrm{HF}}
    +
    \sum_{u=1}^3\braket{\tbar{u}^{[2]}|
    \frac{1}{2}
    \commutator{\commutator{\Phi}{T^{[1]}}}{T^{[1]}}
    | \mathrm{HF}} \\
    &+ {}^t\p^{[2]}\s_{\mathrm{N}}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[2]})
    + {}^t\p^{[1]}\s_{\mathrm{N}}^{[4]}(\tamp{}^{[2]}, \tbar{}^{[2]}) \\
    &= E^{[5]}
    + {}^t\p^{[2]}
    \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
    \commutator{\s}{T^{[2]}} | \mathrm{HF}} \\
    &+ {}^t\p^{[2]}
    \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
    \frac{1}{2}\commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
    | \mathrm{HF}}
    + {}^t\p^{[2]}
    \sum_{u=1}^3\braket{ \tbar{u}^{[2]}
                                 | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
    \\
    &+ {}^t\p^{[1]}
    \sum_{u=1}^3\braket{ \tbar{u}^{[2]}
    \commutator{\s}{T^{[2]}} | \mathrm{HF}}
    + {}^t\p^{[1]}
    \sum_{u=1}^3\braket{ \tbar{u}^{[2]}
    \frac{1}{2}\commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
    | \mathrm{HF}} \\
    &+ {}^t\p^{[1]}
    \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
    \commutator{\commutator{\s}{T^{[1]}}}{T^{[2]}}
    | \mathrm{HF}}
    \\
  &= E^{[5]}
    + {}^t\p^{[2]}\braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} +
    T_2^{[2]}} | \mathrm{HF}} \\
    &+ \highlight{PMS1797}{
    {}^t\p^{[2]}\braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
    } \\
    &+ {}^t\p^{[2]}\braket{\tbar{1}^{[2]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
    + {}^t\p^{[2]}
    \braket{\tbar{1}^{[1]} |
    \frac{1}{2}\commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
    | \mathrm{HF}} \\
    &+ {}^t\p^{[2]}
    \braket{\tbar{2}^{[1]} |
    \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[1]}}
    | \mathrm{HF}}
    +
    {}^t\p^{[1]}
    \braket{\tbar{1}^{[2]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} |
    \mathrm{HF}} \\
    &+
    \highlight{PMS1797}{
    {}^t\p^{[1]}
    \braket{\tbar{2}^{[2]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
    }
    +
    \highlight{PMS1797}{
    {}^t\p^{[1]}
    \braket{\tbar{3}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
    } \\
    &+
    {}^t\p^{[1]}
    \braket{\tbar{1}^{[2]} |
    \frac{1}{2}\commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
    | \mathrm{HF}}
    +
    {}^t\p^{[1]}
    \braket{\tbar{2}^{[2]} |
    \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[1]}}
    | \mathrm{HF}} \\
    &+
    \highlight{PMS1797}{
    {}^t\p^{[1]}
    \braket{\tbar{3}^{[2]} |
    \frac{1}{2}\commutator{\commutator{\s}{T_2^{[1]}}}{T_2^{[1]}}
    | \mathrm{HF}}
    } \\
    &+ {}^t\p^{[1]}
    \braket{\tbar{1}^{[1]} |
    \commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[2]}}
    | \mathrm{HF}}
    + {}^t\p^{[1]}
    \braket{\tbar{2}^{[1]} |
    \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[2]}}
    | \mathrm{HF}} \\
    &+ {}^t\p^{[1]}
    \braket{\tbar{2}^{[1]} |
    \commutator{\commutator{\s}{T_2^{[1]}}}{T_1^{[2]}}
    | \mathrm{HF}}
  \end{aligned}
\end{equation}
Similarly to the exposition in Chapter \ref{ch:QM}, these expressions
will form the basis for the development of methods for the approximate
beyond \acrshort*{CCSD} including connected triples.

@@

In the \acrshort{PTES} scheme, we replace the expectation value of the
source with its \acrshort{CCS} counterpart. The singles amplitudes equation
thus contains a quantum/classical coupling term. However, to avoid
introducing a coupling with the multipliers equations, an additional
approximation is introduced in the polarization equation, where the
source term is truncated to its multiplier-independent part.
As already noted, this leads to the use of \emph{two} different
polarization equations when solving for the amplitudes and the
multipliers.
Within the \acrshort{CC2} model we have the following amplitudes
equations:
\begin{subequations}\label{eq:ptes-cc2-amp}
  \begin{align}
   \denom{\mu_1}\tamp{\mu_1} &+ \braket{\mu_1 | \check{\Phi} +
   \BCHfirst{\check{\Phi}}{Q_2} | \mathrm{HF}}
   + {}^t\p\braket{\mu_1 | \check{\s} | \mathrm{HF}}
             = 0 \label{eq:ptes-cc2-singles-amplitudes} \\
   \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 | \check{\Phi} | \mathrm{HF}}
             = 0 \label{eq:ptes-cc2-doubles-amplitudes} \\
    \V\p &+
    \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}} = 0
  \end{align}
\end{subequations}
For the \acrshort{CC2} multipliers one has:
\begin{subequations}\label{eq:ptes-cc2-mult}
  \begin{align}
    \denom{\mu_1}\tbar{\mu_1} &+
    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}} +
    \braket{\tbar{1} |
    \commutator{\check{\Phi}}{\cluster{\mu_1}}
    + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{Q_2}
    | \mathrm{HF}}
    \nonumber \\
    &+ \braket{\tbar{2} |
    \commutator{\check{\Phi}}{\cluster{\mu_1}}
    | \mathrm{HF}}
    \nonumber \\
    &+
    {}^t\p\braket{\mathrm{HF} | \commutator{\check{\s}_\mathrm{N}}{\cluster{\mu_1}} | \mathrm{HF}}
    +
    {}^t\p\braket{\tbar{1} |
    \commutator{\check{\s}}{\cluster{\mu_1}} | \mathrm{HF}}
             = 0 \label{eq:ptes-cc2-singles-multipliers} \\
    \denom{\mu_2}\tbar{\mu_2} &+
    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}} +
    \braket{\tbar{1} |
    \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}}
             = 0 \label{eq:ptes-cc2-doubles-multipliers} \\
    \V\p &+ \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS} = 0
    \label{eq:ptes-cc2-polarization-mult}
  \end{align}
\end{subequations}
We note that the \acrshort{PTES} approximation is crucial
in the implementation of \acrshort{PE}-\acrshort{CC2} presented by
\citeauthor{Schwabe2012-cf} in \noparcite[ref.][]{Schwabe2012-cf}
The resolution of the identity can in fact still be applied when
assuming the \acrshort{PTES} approximation.
Moreover, the simplified \acrshort{PE}-\acrshort{CC}
(s\acrshort*{PE}\acrshort*{CC}) Lagrangian of \citeauthor{Krause2016-ee}
is essentially the \acrshort{PTE(S)} approximation for
\acrshort{PE}-\acrshort{CC}.\autocite{Krause2016-ee}

>>>>

\subsection*{Notation}

For ease of reference, I will repeat some of the notational conventions
introduced in Chapter \ref{ch:QM} here.
The cluster operator at truncation level $\mathcal{M}$ is:
\begin{equation}
  T = \sum_{u=1}^\mathcal{M} T_u =
  \sum_{u=1}^\mathcal{M}\sum_{\mu_u}\tamp{\mu_u}\cluster{\mu_u},
  \tag{\ref{eq:cluster} from Chapter \ref{ch:QM}}
\end{equation}
where $\tamp{\mu_u}$ and $\cluster{\mu_u}$ are the cluster amplitudes
and cluster operators for the $\mu_u$-th excited determinant at the
$u$-th excitation level, respectively.
The similarity transformed counterpart of the operator $O$ is denoted
by $\overline{O}$:
\begin{equation}
  \overline{O} =\expo{-T}O\expo{T}.
  \tag{\ref{eq:similarity-transformation} from Chapter \ref{ch:QM}}
\end{equation}
We recall that hermiticity of $O$ is not preserved by the similarity
transformation.
Manipulation of similarity transformed operators involves the use of
their \acrlong*{BCH} expansions:
\begin{equation}
  \overline{O}
  =
  O + \BCHfirst{O}{T} + \BCHsecond{O}{T} + \ldots
  =
  O + \tilde{O},
  \tag{\ref{eq:bch} from Chapter \ref{ch:QM}}
\end{equation}
since we are concerned with one- and two-electron operators only,
such expansions will truncate after the third and fifth terms,
respectively.\autocite{Helgaker2000-tz}
The similarity transformation by means of the singles cluster operator
only, \ie the $T_1$-transformation, will be denoted by a "check" on top
of the operator:
\begin{equation}
  \check{O} = \expo{-T_1}O\expo{T_1}.
  \tag{\ref{eq:t1-transformation} from Chapter \ref{ch:QM}}
\end{equation}
The $T_1$-transformation preserves the particle rank of the operator
$O$, see Appendix \ref{app:mathematical-results} for details.
Finally, the $u$-th excited component of the \acrlong*{CC} left state
will be denoted as:
\begin{equation}
  \bra{\tbar{u}} = \sum_{\mu_u}\tbar{\mu_u}\bra{\mu_u},
  \tag{\ref{eq:left-state} from Chapter \ref{ch:QM}}
\end{equation}
where $\tbar{\mu_u}$ are Lagrange multipliers.

@@

It is thus convenient to devise approximations that are able to simplify the
Lagrangian and the governing equations by breaking or weakening the coupling.
While Eq. \eqref{eq:pted-cc-polarization} is \emph{directly} coupled to
the multipliers equation, since the source terms directly depends on
amplitudes and multipliers, it is only \emph{indirectly} coupled to the
amplitudes equation, where only the polarization enters.
Moreover, the leading terms in the correlated source operator
expectation value prominently involve the
singles cluster operator and the singles multiplier state.
Three approximations have currently been proposed: \gls{PTES}, \gls{PTE(S)}
and \gls{PTE}.

@@

To break the coupling, we split the system of equations

@@

term. This adds another $O(N^5)$ step to the iterative

@@

The \acrshort{PTES} approximation can be brought one step further, by
truncating the \acrshort{CCS} expectation value to its multiplier-independent
term already in the effective Lagrangian.
The coupling between the amplitudes and multipliers is broken, while
still improving upon the reference reaction field with correlated
contributions.\autocite{Caricato2011-tx}
The effective Lagrangian would then read as:
\begin{equation}\label{eq:pte-s-cc}
    {}^{\mathrm{PTE(S)}}\lag(\tamp{}, \tbar{}, \p)_\mathcal{M} =
  {}^\mathrm{PTE}\lag(\tamp{}, \tbar{})_\mathcal{M}
  + \frac{1}{2}{}^t\p\V\p + {}^t\p\braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}}
\end{equation}
which uncouples all equations:
\begin{subequations}
  \begin{align}
  {}^{\mathrm{PTE(S)}}\tampEq{\mu_q}(\tamp{}, \tbar{})  &=
   \denom{\mu_q}\tamp{\mu_q} + \braket{\mu_q | \overline{\Phi} | \mathrm{HF}}
   = 0 \\
   {}^{\mathrm{PTE(S)}}\tbarEq{\mu_q}(\tamp{}, \tbar{}, \p) &=
    \denom{\mu_q}\tbar{\mu_q} +
    \braket{\mathrm{HF} | \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}} +
    \sum_{u=1}^{\mathcal{M}}\braket{\tbar{u} |
    \commutator{\overline{\Phi}}{\cluster{\mu_q}} | \mathrm{HF}}
    \nonumber \\
    &+
    {}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}
    = 0 \\
    {}^{\mathrm{PTE(S)}}\Omega_\p(\tamp{}, \tbar{}, \p)
    &=
    \V\p + \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}} = 0
    \label{eq:pte-s-polarization}
  \end{align}
\end{subequations}
The amplitude equations are formally equivalent to those derived in the
\acrshort{PTE} scheme, \emph{vide infra}.
The equations for the multipliers are also formally equivalent to the
\acrshort{PTE} equations, apart from the equation determining the
singles multipliers which features an additional term involving the
polarization:
${}^t\p\braket{\mathrm{HF} | \commutator{\s}{\cluster{\mu_q}} | \mathrm{HF}}\delta_{\mu_q\mu_1}$.
This term is not, however, coupled to the other equations given that
$\p$ is fixed once the singles amplitudes have been determined.

@@

Eventually, the fifth order energy correction is formed including second order
amplitudes, multipliers and polarization:
\begin{equation}\label{eq:pted-mbpt-5th}
  \begin{aligned}
    G^{[5]} &=
    \braket{\mathrm{HF} | \frac{1}{2}
    \commutator{\commutator{\Phi}{T^{[2]}}}{T^{[2]}}
    | \mathrm{HF}}
    +
    \sum_{u=1}^2\braket{\tbar{u}^{[1]}|
    \commutator{\commutator{\Phi}{T^{[2]}}}{T^{[1]}}
    | \mathrm{HF}}
    \\
    &+
    \sum_{u=1}^3\braket{\tbar{u}^{[2]}|
      \commutator{\Phi}{T^{[2]}}
    | \mathrm{HF}}
    +
    \sum_{u=1}^3\braket{\tbar{u}^{[2]}|
    \frac{1}{2}
    \commutator{\commutator{\Phi}{T^{[1]}}}{T^{[1]}}
    | \mathrm{HF}} \\
    &+ {}^t\p^{[2]}\s_{\mathrm{N}}^{[3]}(\tamp{}^{[2]}, \tbar{}^{[2]})
    + {}^t\p^{[1]}\s_{\mathrm{N}}^{[4]}(\tamp{}^{[2]}, \tbar{}^{[2]}) \\
    &= E^{[5]}
    + {}^t\p^{[2]}
    \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
    \commutator{\s}{T^{[2]}} | \mathrm{HF}} \\
    &+ {}^t\p^{[2]}
    \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
    \frac{1}{2}\commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
    | \mathrm{HF}}
    + {}^t\p^{[2]}
    \sum_{u=1}^3\braket{ \tbar{u}^{[2]}
                                 | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
    \\
    &+ {}^t\p^{[1]}
    \sum_{u=1}^3\braket{ \tbar{u}^{[2]}
    \commutator{\s}{T^{[2]}} | \mathrm{HF}}
    + {}^t\p^{[1]}
    \sum_{u=1}^3\braket{ \tbar{u}^{[2]}
    \frac{1}{2}\commutator{\commutator{\s}{T^{[1]}}}{T^{[1]}}
    | \mathrm{HF}} \\
    &+ {}^t\p^{[1]}
    \sum_{u=1}^2\braket{\tbar{u}^{[1]} |
    \commutator{\commutator{\s}{T^{[1]}}}{T^{[2]}}
    | \mathrm{HF}}
    \\
  &= E^{[5]}
    + {}^t\p^{[2]}\braket{\tbar{1}^{[1]} | \commutator{\s}{T_1^{[2]} +
    T_2^{[2]}} | \mathrm{HF}} \\
    &+ \highlight{PMS1797}{
    {}^t\p^{[2]}\braket{\tbar{2}^{[1]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
    } \\
    &+ {}^t\p^{[2]}\braket{\tbar{1}^{[2]} | \commutator{\s}{T^{[1]}} | \mathrm{HF}}
    + {}^t\p^{[2]}
    \braket{\tbar{1}^{[1]} |
    \frac{1}{2}\commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
    | \mathrm{HF}} \\
    &+ {}^t\p^{[2]}
    \braket{\tbar{2}^{[1]} |
    \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[1]}}
    | \mathrm{HF}}
    +
    {}^t\p^{[1]}
    \braket{\tbar{1}^{[2]} | \commutator{\s}{T_1^{[2]} + T_2^{[2]}} |
    \mathrm{HF}} \\
    &+
    \highlight{PMS1797}{
    {}^t\p^{[1]}
    \braket{\tbar{2}^{[2]} | \commutator{\s}{T_2^{[2]} + T_3^{[2]}} | \mathrm{HF}}
    }
    +
    \highlight{PMS1797}{
    {}^t\p^{[1]}
    \braket{\tbar{3}^{[2]} | \commutator{\s}{T_3^{[2]}} | \mathrm{HF}}
    } \\
    &+
    {}^t\p^{[1]}
    \braket{\tbar{1}^{[2]} |
    \frac{1}{2}\commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[1]}}
    | \mathrm{HF}}
    +
    {}^t\p^{[1]}
    \braket{\tbar{2}^{[2]} |
    \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[1]}}
    | \mathrm{HF}} \\
    &+
    \highlight{PMS1797}{
    {}^t\p^{[1]}
    \braket{\tbar{3}^{[2]} |
    \frac{1}{2}\commutator{\commutator{\s}{T_2^{[1]}}}{T_2^{[1]}}
    | \mathrm{HF}}
    } \\
    &+ {}^t\p^{[1]}
    \braket{\tbar{1}^{[1]} |
    \commutator{\commutator{\s}{T_1^{[1]}}}{T_1^{[2]}}
    | \mathrm{HF}}
    + {}^t\p^{[1]}
    \braket{\tbar{2}^{[1]} |
    \commutator{\commutator{\s}{T_1^{[1]}}}{T_2^{[2]}}
    | \mathrm{HF}} \\
    &+ {}^t\p^{[1]}
    \braket{\tbar{2}^{[1]} |
    \commutator{\commutator{\s}{T_2^{[1]}}}{T_1^{[2]}}
    | \mathrm{HF}}
  \end{aligned}
\end{equation}
Similarly to the exposition in Chapter \ref{ch:QM}, these expressions
will form the basis for the development of approximations
beyond \acrshort*{CCSD} including connected triples.

@@

In the \acrshort{PTES} scheme, we replace the expectation value of the
source with its \acrshort{CCS} counterpart. The singles amplitudes equation
thus contains a quantum/classical coupling term. However, to avoid
the coupling with the multipliers equations, an additional
approximation is introduced in the polarization equation, where the
source term is truncated to its multiplier-independent part.
As already noted, this leads to the use of \emph{two} different
polarization equations when solving for the amplitudes and the
multipliers.
Within the \acrshort{CC2} model we have the following amplitudes
equations:
\begin{subequations}\label{eq:ptes-cc2-amp}
  \begin{align}
   \denom{\mu_1}\tamp{\mu_1} &+ \braket{\mu_1 | \check{\Phi} +
   \BCHfirst{\check{\Phi}}{Q_2} | \mathrm{HF}}
   + {}^t\p\braket{\mu_1 | \check{\s} | \mathrm{HF}}
             = 0 \label{eq:ptes-cc2-singles-amplitudes} \\
   \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 | \check{\Phi} | \mathrm{HF}}
             = 0 \label{eq:ptes-cc2-doubles-amplitudes} \\
    \V\p &+
    \braket{\mathrm{HF} | \check{\s}_{\mathrm{N}} | \mathrm{HF}} = 0
  \end{align}
\end{subequations}
For the \acrshort{CC2} multipliers one has:
\begin{subequations}\label{eq:ptes-cc2-mult}
  \begin{align}
    \denom{\mu_1}\tbar{\mu_1} &+
    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}} +
    \braket{\tbar{1} |
    \commutator{\check{\Phi}}{\cluster{\mu_1}}
    + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{Q_2}
    | \mathrm{HF}}
    \nonumber \\
    &+ \braket{\tbar{2} |
    \commutator{\check{\Phi}}{\cluster{\mu_1}}
    | \mathrm{HF}}
    \nonumber \\
    &+
    {}^t\p\braket{\mathrm{HF} | \commutator{\check{\s}_\mathrm{N}}{\cluster{\mu_1}} | \mathrm{HF}}
    +
    {}^t\p\braket{\tbar{1} |
    \commutator{\check{\s}}{\cluster{\mu_1}} | \mathrm{HF}}
             = 0 \label{eq:ptes-cc2-singles-multipliers} \\
    \denom{\mu_2}\tbar{\mu_2} &+
    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}} +
    \braket{\tbar{1} |
    \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}}
             = 0 \label{eq:ptes-cc2-doubles-multipliers} \\
    \V\p &+ \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS} = 0
    \label{eq:ptes-cc2-polarization-mult}
  \end{align}
\end{subequations}
We note that the \acrshort{PTES} approximation is crucial
in the implementation of \acrshort{PE}-\acrshort{CC2} presented by
\citeauthor{Schwabe2012-cf} in \noparcite[ref.][]{Schwabe2012-cf}
The resolution of the identity can in fact still be applied when
assuming the \acrshort{PTES} approximation.
Moreover, the simplified \acrshort{PE}-\acrshort{CC}
(s\acrshort*{PE}\acrshort*{CC}) Lagrangian of \citeauthor{Krause2016-ee}
is essentially the \acrshort{PTE(S)} approximation for
\acrshort{PE}-\acrshort{CC}.\autocite{Krause2016-ee}
